|

### 2023-05-29 manuscript/manuscript_1.0.Rmd
New version of the manuacrip. New files to manage the authors affiliations

---
title: |
  The role of learning in status signalling: a modeling approach
author:
- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
- C. Daniel Cadena*
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex # default  # word_document: default
fig_caption: yes
fontsize: 12pt
header-includes: |
  ```{=latex}
    \usepackage{float}
    \usepackage{unicode-math}
    \usepackage{setspace}\doublespacing
    \usepackage[backref=page]{hyperref}
    \linespread{2}
    \usepackage{lineno}
    \linenumbers
    \floatplacement{figure}{H}
    \newcommand{\beginsupplement}{ \setcounter{table}{0}         \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  ```
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: ../SocialMorphCue.bib
spacing: double
style: C:\Users\andre\Zotero\styles\elife.csl
abstract: |
  Adaptive behavioural responses often depend on qualities of the interacting
  partner of an individual. For example, when competing for resources, an individual
  might be better off escalating fights with individuals of lower quality, while
  restraining  from fighting individuals of higher quality. Communication systems
  involving signals of quality allow individuals to reduce uncertainty regarding the  
  fighting ability of their partners and make more adaptive behavioral decisions. 
  However, dishonest individuals can destabilize such communications systems. 
  One open question is whether cognitive mechanisms, such as learning, 
  can maintain the honesty of signals, thus, favoring their evolutionary 
  stability. Here we present evolutionary simulations, where individuals 
  can produce a signal proportional to their quality; and learn along 
  their lifetime the best response to the signal emitted by their peers. In
  these simulations, learning on the receiver side can mediate the evolution of signals
  of quality on the sender side. When the cost of the signal is proportional to the
  quality of the sender populations are only composed of honest signalers. When the
  cost is not proportional to the quality of the signaler, the population is composed
  of both honest and dishonest signalers. We argue that learning can be a general
  cognitive mechanism playing a role in a wide range of communication systems."
---

@@

In dominance and aggressive interactions, a crucial piece of
information to guide individual actions is the fighting ability of interacting 
partners, which is sometimes referred to as Resource Holding Potential 
[RHP, Parker -@parker_Assessment_1974], or simply as quality. 
Responsiveness to the quality of an opponent is central 
to communication systems such as those in which individuals 
exhibit badges of status, where a signal conveys quality 
[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
Such signals must be arbitrary, meaning the signals could potentially be 
produced by high and low quality individuals alike. 
Generally signals can be evolutionarily stable whenever low quality individuals 
pay a higher fitness cost for carrying the signal such that it is not 
in the interest of low quality individuals 
to fake quality by producing a large badge. Signal costs inversely proportional
to quality can come about as a result of the production costs of the signal, 
in which case the signal works like a handicap [@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990; @zahavi_Mate_1975]. 
Alternatively, the costs can come from 
the aggressive reaction of receivers when the convention of the communication 
is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. 
<!-- should this last part of the paragraph stay ?-->
<!-- An alternative system of communication is one where individuals  -->
<!-- gather and store information about each individual they interact with.  -->
<!-- This information should reliably allow each individual to react adaptively  -->
<!-- to the quality difference with their opponent. This type of  -->
<!-- assessment strategy is often referred to as individual  -->
<!-- recognition [@whitfield_Plumage_1986]. Evidently, individual recognition  -->
<!-- is restricted to  certain interaction structures, because individuals  -->
<!-- are limited by cognitive abilities in how many individuals  -->
<!-- they can recognize and track. Thus, individual recognition and  -->
<!-- badges of status have been assumed to be alternative communication  -->
<!-- systems fulfilling the same function, reducing the costs of  -->
<!-- aggression in contests over resources.  -->

@@

Associative learning is a taxonomically widespread cognitive 
mechanism that allows individuals to associate rewards with environmental stimuli 
and thus behave adaptively 
[@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
@behrens_Associative_2008].
Theory has shown that natural selection favours such associations 
in complex environments where conditions are difficult to predict 
[@dridi_Environmental_2016]. Associative learning is a flexible cognitive 
mechanism whose underpinnings show interspecific variation 
[@enquist_Power_2016; @quinones_Reinforcement_2020]. Associative learning 
is not a single mechanisms, but rather a set of cognitive structures 
and processes that can vary in their scope and complexity. Presumably, 
these structures and processes have been modified 
by natural selection and could provide new explanations for behavioural 
variation. Despite all this, associative learning is 
not often included in the narrative of evolutionary explanations of 
behavioural patterns [@fawcett_Exposing_2013; @kamil_Optimal_1983; 
@mcauliffe_Psychology_2015]. Computational models of evolution 
can play an important role to overcome the lack of integration 
between learning and evolution. Reinforcement learning theory 
encompasses a series of computational methods inspired 
on the psychological and neurological mechanisms of associative 
learning [@sutton_Reinforcement_2018]. This set of algorithms allows the 
implementation of biologically realistic problems, 
capturing the essence of learning processes 
[@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
Furthermore, these algorithms can be embedded in evolutionary simulations to 
generate theoretical predictions of the effect that learning can 
have in behavioural evolution [@leimar_Learning_2019;@leimar_Effects_2022].

@@

# Top panels -  A the sender code and  B learning for the receiver code

library(dplyr)
honestBig<-readPNG(here("Images/cart_honestBig.png"))
honestSmall<-readPNG(here("Images/cart_honestSmall.png"))
dishonest<-readPNG(here("Images/cart_dishonest.png"))
par(plt=posPlot(numplotx = 2,idplotx = 1,numploty = 1,idploty = 1,upboundx = 91),
    las=1,mfrow=c(2,1),mgp=c(2,0.7,0))
  rangPars<-data.frame(honest=c(4,2),dishonest=c(0,0))
 rangQual<-seq(0,1,by = 0.01)
 linesSender<-sapply(rangPars,function(x)logist(rangQual,alpha = x[2],beta = x[1]))
 plot(0,0,type="l",xlab="",xlim=c(0,1),
      ylab ="",cex.lab=1,lwd=2,col="grey",cex.axis=1,ylim=c(0,1.1),xaxt="n")
 axis(1)
 text(x = 0.08,y=1.05,"A",cex = 1.5)
 mtext("Quality",1,line = 2,cex = 1,las = 0)
 matlines(x=rangQual,y=linesSender,lwd=4,col=colRuns,lty=1)
 mtext("Badge size",2,line = 2.3,cex = 1,las = 0)
 # include bird cartoons
 rasterImage(honestBig,0.75,0.87,0.95,1.07)
 rasterImage(honestSmall,0.05,0.01,0.25,0.21)
 rasterImage(dishonest,0.75,0.5,0.95,0.7)
 rasterImage(dishonest,0.05,0.5,0.25,0.7)
 # legend("left",legend=rangeBeta,col=colRuns,lwd=3,title = expression(beta),cex=2,bty="n")

@@

  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
                     function(x){x=
                       logist(totRBF(rangx,
                                     centers,sigSquar,
                                     as.double(
                                       tempPop[indId==x,.SD,
                                               .SDcol=grep("WeightAct",
                                                           names(tempPop),
                                                           value = TRUE)
                                               ])),alpha=0,beta = 1)})
  matplot(x=rangx,y=dataIndAct,
          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
            findInterval(tempPop[indId %in%randInds,Quality],
                         colorbreaksQual)],lty = 1,
          cex.axis=cexAxis)
  axis(side = 1,cex.axis=cexAxis,padj = 0)
  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
        col=1,lwd=3,lty=2)
  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
  #      cex=2)
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
                  lowboundx = 8,lowboundy = 15,
                  upboundy = 99), las=1,new=TRUE)

@@

  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
                     function(x){x=
                       logist(totRBF(rangx,
                                     centers,sigSquar,
                                     as.double(
                                       tempPop[indId==x,.SD,
                                               .SDcol=grep("WeightAct",
                                                           names(tempPop),
                                                           value = TRUE)
                                               ])),alpha=0,beta = 1)})
  matplot(x=rangx,y=dataIndAct,
          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
            findInterval(tempPop[indId %in%randInds,Quality],
                         colorbreaksQual)],lty = 1,
          cex.axis=cexAxis)
  axis(side = 1,cex.axis=cexAxis,padj = 0)
  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
        col=1,lwd=3,lty=2)
  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
  #      cex=2)
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
                  lowboundx = 8,lowboundy = 15,
                  upboundy = 99), las=1,new=TRUE)

@@

  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
                     function(x){x=
                       logist(totRBF(rangx,
                                     centers,sigSquar,
                                     as.double(
                                       tempPop[indId==x,.SD,
                                               .SDcol=grep("WeightAct",
                                                           names(tempPop),
                                                           value = TRUE)
                                               ])),alpha=0,beta = 1)})
  matplot(x=rangx,y=dataIndAct,
          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
            findInterval(tempPop[indId %in%randInds,Quality],colorbreaksQual)],lty = 1,
          cex.axis=cexAxis)
  axis(side = 1,cex.axis=cexAxis,padj = 0)
  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
        col=1,lwd=3,lty=2)
  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
  #      cex=2)
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
                  lowboundx = 8,lowboundy = 15,
                  upboundy = 99), las=1,new=TRUE)

@@

### The peaceful, the clever and the aggressive

@@

yaxtAll<-c("s","n","n")
xlabAll<-c("",expression(beta[s]),"")
ylabAll<-c(expression(alpha[s]),"","")
titleAll<-c("Peaceful","Clever","Aggressive")
colorAll<-c("dodgerblue2","green4","#E31A1C")
multDiscrPallet[repsScen+1]<-colorAll
plot.new()
for(PAr in popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)]){
  count<-match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])
  # par(plt=posPlot(numploty = 2,idploty = 2,
  #                 numplotx = length(popOneInd[,unique(get(shortSce))]),
  #                 idplotx = count)+c(0,0,-0.05,-0.05),
  #     las=1,new=TRUE,
  #     yaxt=yaxtAll[match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])],
  #     cex.axis=1.5,new=TRUE)
  # plot(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&seed %in% 12:15],
  #      alpha~beta,ylab="",
  #      xlab="", pch=20,cex.lab=3,cex.axis=2,las=1,cex=2,xaxt="n",
  #      ylim=range(popOneInd[,alpha])*c(1.05,1.05),
  #      xlim=range(popOneInd[,beta]),
  #      # col=alpha(multDiscrPallet[seed+1],0.3))
  #      col=multDiscrPallet[seed+1])
  # # points(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&
  # #                         seed == repsScen[count]],
  # #      alpha~beta,col=multDiscrPallet[seed+1],cex=3,pch=20)
  # lines(x=c(0,0),y=range(popOneInd[,alpha]),col="grey",
  #       lwd=2)
  # lines(y=c(0,0),x=range(popOneInd[,beta]),col="grey",
  #       lwd=2)

@@

Here we presented a model that integrates the basic mechanisms of 
learning with the fundamental rules of evolutionary processes. 
In the model the response of a signal in a 
communication system is mediated by learning processes. Individuals 
throughout their life learn the best way to respond to a quantitative 
trait of their interactive partners. 
The learning process can mediate the evolution 
of an honest signal under the handicap 
principle [@grafen_Biological_1990;@zahavi_Mate_1975]. This conclusion is not 
different from the classical genetic models, where responses to the signal are
inherent to the individual. However, unlike classical genetic models,
our model shows that learning can also mediate a signal polymorphism in the 
absence of the handicap principle. Under this polymorphic equilibrium different 
types of individuals, with honest and dishonest signals, compose the population. 
The amount of information gathered through learning, as a well as the 
initial conditions of the learning process, are crucial for the emergence
and maintenance of this variation. The emergence of the polymorphism requires
a limit in the amount of information collected by individuals, as well as
peaceful behaviour in naive individuals.

Associative learning is a powerful mechanism to learn about the world, about 
social partners and about one's abilities relative to the social context. 
By associating cues and signals with fitness relevant outcomes individuals
collect information that allows them to improve their reproductive potential. 
This is very obvious in classical examples such as when animals learn to avoid 
food that makes them sick, or use environmental cues to find food. 
More recent ideas have highlighted the potential role that associations 
can have in a social contexts such as: hierarchy formation 
[@leimar_Evolution_2021;@leimar_Reproductive_2022;@leimar_Effects_2022],
and cooperation [@leimar_Learning_2019;@dridi_Learning_2018]. In this 
theoretical examples individuals use various sources of reward to make
adaptive choices in a given social context. Here we extend this logic to the 
evolution of a communication system mediated by a quality signal. 
Previous evolutionary models of badges of status assumed that individuals 
responded using a behavioural reaction norm, where the opponent's badge 
and individual's own quality determined the behaviour [@botero_Evolution_2010]. 
However, there isn't a clear mechanism to justify the assumption that 
individuals inherently know their own quality, particularly relative to their peers.
In our model, individuals not only learn about the quality signal, but also 
learn about their own quality relative to others. This can be seen throughout
our simulations in the responses developed by individuals of different quality. 
Because individuals learn about their own quality, 
even in the absence of an honest signal, they are able to make adaptive choices. 
Thus the learning process we have modelled and the response it mediates 
has fitness relevant consequences even in the absence of an honest signal.

>>>>

---
title: |
  The role of learning in status signalling: a modeling approach
author:
- Andrés E. Quiñones:
    email: "andresqp@gmail.com"
    institute: [andes, unine]
    correspondence: true
- Redouan Bshary:
    institute: unine
- C. Daniel Cadena:
    institute: andes
institute:
- andes: Laboratorio de Biología Evolutiva de Vertebrados, Departamento  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia
- unine: Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  # word_document: default
  pdf_document:
    latex_engine: xelatex # default  #
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
fig_caption: yes
fontsize: 12pt
header-includes: |
  ```{=latex}
    \usepackage{float}
    \usepackage{unicode-math}
    \usepackage{setspace}\doublespacing
    \usepackage[backref=page]{hyperref}
    \linespread{2}
    \usepackage{lineno}
    \linenumbers
    \floatplacement{figure}{H}
    \newcommand{\beginsupplement}{ \setcounter{table}{0}         \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  ```
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: ../SocialMorphCue.bib
spacing: double
style: C:\Users\andre\Zotero\styles\elife.csl
abstract: |
  Adaptive behavioural responses often depend on qualities of the interacting
  partner of an individual. For example, when competing for resources, an individual
  might be better off escalating fights with individuals of lower quality, while
  restraining  from fighting individuals of higher quality. Communication systems
  involving signals of quality allow individuals to reduce uncertainty regarding the
  fighting ability of their partners and make more adaptive behavioral decisions. 
  However, dishonest individuals can destabilize such communications systems. 
  One open question is whether cognitive mechanisms, such as learning, 
  can maintain the honesty of signals, thus, favoring their evolutionary 
  stability. Here we present evolutionary simulations, where individuals 
  can produce a signal proportional to their quality; and learn along their
  lifetime the best response to the signal emitted by their peers. Our simulations
  replicate previous results where the handicap principle mediates the evolution 
  of signals. As in traditional models without learning, in our simulations 
  learning on the receiver side can mediate the evolution of signals
  of quality on the sender side. When the cost of the signal is proportional to the
  quality of the sender populations are only composed of honest signalers. When the
  cost is not proportional to the quality of the signaler tranditional models 
  predict the absense of a signal, in contrast learning facilitates the 
  evolution of a polymorphim where population is composed of both honest and 
  dishonest signalers. We argue that learning can be a general cognitive mechanism 
  role wide range of communication systems and more generally
  behavioral responses."
---

@@

In dominance and aggressive interactions, a crucial piece of
information to guide individual actions is the fighting ability of interacting 
partners, which is sometimes referred to as Resource Holding Potential 
[RHP, Parker -@parker_Assessment_1974], or simply as quality. 
Responsiveness to the quality of an opponent is central 
to communication systems such as those in which individuals 
exhibit badges of status, where a signal conveys quality 
[@johnstone_Badges_1993; @rohwer_Social_1975]. 
Such signals must be arbitrary, meaning the signals could potentially be 
produced by high and low quality individuals alike. 
Generally signals can be evolutionarily stable whenever low quality individuals 
pay a higher fitness cost for carrying the signal such that it is not 
in the interest of low quality individuals 
to fake quality by producing a large badge. Signal costs inversely proportional
to quality can come about as a result of the production costs of the signal, 
in which case the signal works like a handicap [@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990; @zahavi_Mate_1975]. 
Alternatively, the costs can come from 
the aggressive reaction of receivers when the convention of the communication 
is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. 
<!-- should this last part of the paragraph stay ?-->
<!-- An alternative system of communication is one where individuals  -->
<!-- gather and store information about each individual they interact with.  -->
<!-- This information should reliably allow each individual to react adaptively  -->
<!-- to the quality difference with their opponent. This type of  -->
<!-- assessment strategy is often referred to as individual  -->
<!-- recognition [@whitfield_Plumage_1986]. Evidently, individual recognition  -->
<!-- is restricted to  certain interaction structures, because individuals  -->
<!-- are limited by cognitive abilities in how many individuals  -->
<!-- they can recognize and track. Thus, individual recognition and  -->
<!-- badges of status have been assumed to be alternative communication  -->
<!-- systems fulfilling the same function, reducing the costs of  -->
<!-- aggression in contests over resources.  -->

@@

Associative learning is a taxonomically widespread cognitive 
mechanism that allows individuals to associate rewards with environmental stimuli 
and thus behave adaptively 
[@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
@behrens_Associative_2008].
Theory has shown that natural selection favours such associations 
in complex environments where conditions are difficult to predict 
[@dridi_Environmental_2016]. Associative learning is a flexible cognitive 
mechanism whose underpinnings show interspecific variation 
[@enquist_Power_2016; @quinones_Reinforcement_2020]. Associative learning 
is not a single mechanisms, but rather a set of cognitive structures 
and processes that can vary in their scope and complexity. Presumably, 
these structures and processes have been modified 
by natural selection and could provide new explanations for behavioural 
variation. Despite all this, associative learning is 
not often included in the narrative of evolutionary explanations of 
behavioural patterns [@fawcett_Exposing_2013; @kamil_Optimal_1983; 
@mcauliffe_Psychology_2015]. Computational models of evolution 
can play an important role to overcome the lack of integration 
between learning and evolution. Reinforcement learning theory 
encompasses a series of computational methods inspired 
on the psychological and neurological mechanisms of associative 
learning [@sutton_Reinforcement_2018]. This set of algorithms allows the 
implementation of biologically realistic problems, 
capturing the essence of learning processes 
[@frankenhuis_Enriching_2018; @quinones_Reinforcement_2020]. 
Furthermore, these algorithms can be embedded in evolutionary simulations to 
generate theoretical predictions of the effect that learning can 
have in behavioural evolution [@leimar_Learning_2019;@leimar_Effects_2022].

@@

# Top panels -  A the sender code and  B learning for the receiver code
# load(here("manuscript","manuscript_1.0_cache","latex","model-struc_6905f4797a1faa21041511a8dcdf1eb9.RData"))
library(dplyr)
honestBig<-readPNG(here("Images/cart_honestBig.png"))
honestSmall<-readPNG(here("Images/cart_honestSmall.png"))
dishonest<-readPNG(here("Images/cart_dishonest.png"))
par(plt=posPlot(numplotx = 2,idplotx = 1,numploty = 1,idploty = 1,upboundx = 91),
    las=1,mfrow=c(2,1),mgp=c(2,0.7,0))
  rangPars<-data.frame(honest=c(4,2),dishonest=c(0,0))
 rangQual<-seq(0,1,by = 0.01)
 linesSender<-sapply(rangPars,function(x)logist(rangQual,alpha = x[2],beta = x[1]))
 plot(0,0,type="l",xlab="",xlim=c(0,1),
      ylab ="",cex.lab=1,lwd=2,col="grey",cex.axis=1,ylim=c(0,1.1),xaxt="n")
 axis(1)
 text(x = 0.08,y=1.05,"A",cex = 1.5)
 mtext("Quality",1,line = 2,cex = 1,las = 0)
 matlines(x=rangQual,y=linesSender,lwd=4,col=colRuns,lty=1)
 mtext("Badge size",2,line = 2.3,cex = 1,las = 0)
 # include bird cartoons
 rasterImage(honestBig,0.75,0.87,0.95,1.07)
 rasterImage(honestSmall,0.05,0.01,0.25,0.21)
 rasterImage(dishonest,0.75,0.5,0.95,0.7)
 rasterImage(dishonest,0.05,0.5,0.25,0.7)
 # legend("left",legend=rangeBeta,col=colRuns,lwd=3,title = expression(beta),cex=2,bty="n")

@@

  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
                     function(x){x=
                       logist(totRBF(rangx,
                                     centers,sigSquar,
                                     as.double(
                                       tempPop[indId==x,.SD,
                                               .SDcol=grep("WeightAct",
                                                           names(tempPop),
                                                           value = TRUE)
                                               ])),alpha=0,beta = 1)})
  matplot(x=rangx,y=dataIndAct,
          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
            findInterval(tempPop[indId %in%randInds,Quality],
                         colorbreaksQual)],lty = 1,
          cex.axis=cexAxis)
  axis(side = 1,cex.axis=cexAxis,padj = 0)
  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
        col=1,lwd=3,lty=2)
  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
  if(count==1){
   color.bar.aeqp(paletteMeans(100),min =min(colorbreaksQual),
               max = max(colorbreaksQual),nticks = 3,
               cex.tit = 0.9,title = "",
               numplotx = 20,numploty = 14,idplotx =3,idploty = 1) 
  }
  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
  #      cex=2)
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
                  lowboundx = 8,lowboundy = 15,
                  upboundy = 99), las=1,new=TRUE)

@@

  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
                     function(x){x=
                       logist(totRBF(rangx,
                                     centers,sigSquar,
                                     as.double(
                                       tempPop[indId==x,.SD,
                                               .SDcol=grep("WeightAct",
                                                           names(tempPop),
                                                           value = TRUE)
                                               ])),alpha=0,beta = 1)})
  matplot(x=rangx,y=dataIndAct,
          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
            findInterval(tempPop[indId %in%randInds,Quality],
                         colorbreaksQual)],lty = 1,
          cex.axis=cexAxis)
  axis(side = 1,cex.axis=cexAxis,padj = 0)
  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
        col=1,lwd=3,lty=2)
  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
  #      cex=2)
  if(count==1){
   color.bar.aeqp(paletteMeans(100),min =min(colorbreaksQual),
               max = max(colorbreaksQual),nticks = 3,
               cex.tit = 0.9,title = "",
               numplotx = 20,numploty = 14,idplotx =3,idploty = 1) 
  }
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
                  lowboundx = 8,lowboundy = 15,
                  upboundy = 99), las=1,new=TRUE)

@@

  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
                     function(x){x=
                       logist(totRBF(rangx,
                                     centers,sigSquar,
                                     as.double(
                                       tempPop[indId==x,.SD,
                                               .SDcol=grep("WeightAct",
                                                           names(tempPop),
                                                           value = TRUE)
                                               ])),alpha=0,beta = 1)})
  matplot(x=rangx,y=dataIndAct,
          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
            findInterval(tempPop[indId %in%randInds,Quality],colorbreaksQual)],lty = 1,
          cex.axis=cexAxis)
  axis(side = 1,cex.axis=cexAxis,padj = 0)
  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
        col=1,lwd=3,lty=2)
  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
  #      cex=2)
  if(count==1){
   color.bar.aeqp(paletteMeans(100),min =min(colorbreaksQual),
               max = max(colorbreaksQual),nticks = 3,
               cex.tit = 0.9,title = "",
               numplotx = 20,numploty = 14,idplotx =3,idploty = 1) 
  }
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
                  lowboundx = 8,lowboundy = 15,
                  upboundy = 99), las=1,new=TRUE)

@@

### The peaceful, the aggressive and the clever

@@

yaxtAll<-c("s","n","n")
xlabAll<-c("",expression(beta[s]),"")
ylabAll<-c(expression(alpha[s]),"","")
titleAll<-c("Peaceful","Aggressive","Clever")
colorAll<-c("dodgerblue2","#E31A1C","green4")
multDiscrPallet[repsScen+1]<-colorAll
plot.new()
for(PAr in popOneInd[,unique(get(shortSce))][c(3,2,1)]){
  count<-match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])
  # par(plt=posPlot(numploty = 2,idploty = 2,
  #                 numplotx = length(popOneInd[,unique(get(shortSce))]),
  #                 idplotx = count)+c(0,0,-0.05,-0.05),
  #     las=1,new=TRUE,
  #     yaxt=yaxtAll[match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])],
  #     cex.axis=1.5,new=TRUE)
  # plot(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&seed %in% 12:15],
  #      alpha~beta,ylab="",
  #      xlab="", pch=20,cex.lab=3,cex.axis=2,las=1,cex=2,xaxt="n",
  #      ylim=range(popOneInd[,alpha])*c(1.05,1.05),
  #      xlim=range(popOneInd[,beta]),
  #      # col=alpha(multDiscrPallet[seed+1],0.3))
  #      col=multDiscrPallet[seed+1])
  # # points(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&
  # #                         seed == repsScen[count]],
  # #      alpha~beta,col=multDiscrPallet[seed+1],cex=3,pch=20)
  # lines(x=c(0,0),y=range(popOneInd[,alpha]),col="grey",
  #       lwd=2)
  # lines(y=c(0,0),x=range(popOneInd[,beta]),col="grey",
  #       lwd=2)

@@

Here we presented a model that integrates the basic mechanisms of 
learning with the fundamental rules of evolutionary processes. 
In the model the response of a signal in a 
communication system is mediated by learning processes. Individuals 
throughout their life learn the best way to respond to a quantitative 
trait of their interactive partners. 
The learning process can mediate the evolution 
of an honest signal under the handicap 
principle [@grafen_Biological_1990;@zahavi_Mate_1975]. This conclusion is not 
different from the classical genetic models, where responses to the signal are
inherent to the individual. However, unlike classical genetic models,
our model shows that learning can also mediate a signal polymorphism in the 
absence of the handicap principle. Under this polymorphic equilibrium different 
types of individuals, with honest and dishonest signals, compose the population. 
The amount of information gathered through learning, as a well as the 
initial conditions of the learning process, are crucial for the emergence
and maintenance of this variation. The emergence of the polymorphism requires
a limit in the amount of information collected by individuals, as well as
peaceful behaviour in naive individuals. Furthermore, we have shown that those 
conditions can be reached when evolutionary processes drive changes in the learning
system. Thus, the diversification process we present might play a role in natural
populations.

Associative learning is a powerful mechanism to learn about the world, about 
social partners and about one's abilities relative to the social context. 
By associating cues and signals with fitness relevant outcomes individuals
collect information that allows them to improve their reproductive potential. 
This is very obvious in classical examples such as when animals learn to avoid 
food that makes them sick, or use environmental cues to find food. 
More recent ideas have highlighted the potential role that associations 
can have in a social contexts such as: hierarchy formation 
[@leimar_Evolution_2021;@leimar_Reproductive_2022;@leimar_Effects_2022],
and cooperation [@leimar_Learning_2019;@dridi_Learning_2018]. In this 
theoretical examples individuals use various sources of reward to make
adaptive choices in a given social context. Here we extend this logic to the 
evolution of a communication system mediated by a quality signal. 
Previous evolutionary models of badges of status assumed that individuals 
responded using a behavioural reaction norm, where the opponent's badge 
and individual's own quality determined the behaviour [@botero_Evolution_2010]. 
However, there isn't a clear mechanism to justify the assumption that 
individuals inherently know their own quality, particularly relative to their peers.
In our model, individuals not only learn about the quality signal, but also 
learn about their own quality relative to others. This can be seen throughout
our simulations in the responses developed by individuals of different quality. 
Because individuals learn about their own quality, 
even in the absence of an honest signal, they are able to make more 
adaptive choices. Thus the learning process we have modelled and the 
response it mediates has fitness relevant consequences even in the 
absence of an honest signal. This is confirmed by simulation where 
the speed of learning is allowed to change subject to evolutionary processes. 
In those simulations, natural selection always
maintained learning rates above zero. Our simulations however, 
only captured the effect of selection mediated by the social 
game on the learning parameters. Inter-specific variation in cognitive 
abilities mediated by environmental differences has been reported elsewhere
[@sonnenberg_Natural_2019]. If such variation is partly mediated by changes 
that affect general cognitive processes, they could impact the outcome 
of communication systems like the one we modelled.
